All Samples Model

### 2018-03-12 AllSamplesModel.Rmd
...

globalCNdf$Status = factor(globalCNdf$Status, levels=sort(unique(globalCNdf$Status), decreasing=T), ordered=T)
globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
cnLegend = get.legend(globalCN.plot)
#globalCN.plot = globalCN.plot + theme(legend.position = 'none')

## NDBE
nb.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')', sep=''))
globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')

globalCN.plot = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
ggsave(filename='plots/globalCN.png', plot=globalCN.plot, width=16, height=4.6, units='in', dpi=300)
globalCN.plot

globalBECN.plot = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") 
ggsave(filename='plots/globalBECN.png', plot=globalBECN.plot, width=16, height=4.6, units='in', dpi=300)
globalBECN.plot

@@

```{r noLGD, echo=F, warning=F, message=F, fig.width=6, fig.height=6, eval=F}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

@@

chh$ymax = ifelse(chh$coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value))
chh$gl = factor(ifelse(chh$coef > 0, 'gain', 'loss'))

@@

ggsave(filename='plots/discovery_roc.png', plot=roc.plot(roc), width=6, height=4, units='in', dpi=300)

@@

ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)

@@

ggsave(filename='plots/pred_calib.png', plot=pcal, width=3, height=3, units='in', dpi=300)

@@

unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
  #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
unadjRR
ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)

@@

@@

ggsave(filename='plots/pred_by_path.png', plot=rp, width=8, height=8, units='in', dpi=300)

@@

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='orangered4', show.legend = F) + 
  geom_errorbar(color='grey21', width=0.2) +
  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
rp
ggsave(filename='plots/pred_endo.png', plot=rp, width=9, height=7, units='in', dpi=300)

#endos$months.before.final = endos$months.before.final*-1
#me = endos %>% group_by(Patient) %>% dplyr::summarise('max.endo' = max(i), 'months' = min(months.before.final))
```
# TODO number endoscopies so the images are scaled 
`r pander(as.data.frame(pred.confidence))`

@@

genes = cbind(subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer) & Mutation.Types != 'T', select=c('Gene.Symbol','Genome.Location', colnames(loc))), 'value'=min(globalCNdf$mean.value)+2*sd(globalCNdf$mean.value), 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))

@@

ggsave(filename='plots/hazards/chr17.png', plot=g17, width=10, height=5, units='in', dpi=300)

@@

```{r echo=F, eval=T}
dirA = '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent'
dirB = '~/Data/Ellie/Analysis/downsampled5G_NDBE'

@@

predA = list.files(dirA, 'predictions_all_exAHM0320.Rdata', full.names = T)
predB = list.files(dirB, 'predictions_all_exAHM0320.Rdata', full.names = T)

files = c()
if (length(predA) < 1)
  files = c(list.files(dirA, pattern='raw', full.names=T))
if (length(predB) < 1)
  files = c(files, list.files(dirB, pattern='raw', full.names=T))

if (length(files) > 0) {
  sampleNames = sub('_raw.txt', '', basename(files))
  dspred = data.frame(matrix(ncol=2, nrow=length(sampleNames), dimnames=list(sampleNames,c('Prob','RR'))))

  predA = sub('_raw.txt','',basename(c(list.files(dirA, pattern='raw', full.names=T))))
  predB = sub('_raw.txt','',basename(c(list.files(dirB, pattern='raw', full.names=T))))

  for (n in 1:length(files)) {
    print(n)
    print(sampleNames[n])

@@

    dspred[n,] = c(preds[,1],RR[,1])
    print(dspred[1:n,])
  }
    dspred$Type = NA
    dspred[predA,'Type'] = 'AdjacentBE'
    dspred[predB,'Type'] = 'BEST2 NDBO'
    pred = dspred

    dspred = subset(pred, Type == 'AdjacentBE')

    save(dspred, file=paste(dirA, 'predictions_all_exAHM0320.Rdata', sep='/'))

    dspred = subset(pred, Type == 'BEST2 NDBO')
    save(dspred, file=paste(dirB, 'predictions_all_exAHM0320.Rdata', sep='/'))

    dspred = pred

} else {
  load(predB, verbose=T)
  dspredB = dspred

  load(predA, verbose=T)

  dspred = rbind(dspred, dspredB)
  dspred$Type = NA

  dspred[sub('_raw.*','',basename(list.files(dirA, pattern='raw', full.names=T))),'Type'] = 'AdjacentBE'
  dspred[sub('_raw.*','',basename(list.files(dirB, pattern='raw', full.names=T))),'Type'] = 'BEST2 NDBO'

>>>>

globalCNdf$Status = factor(globalCNdf$Status, levels=sort(unique(globalCNdf$Status), decreasing=T), ordered=T)
globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value)) 
cnLegend = get.legend(globalCN.plot + theme(legend.position='bottom'))

globalCN.plot = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
ggsave(filename='plots/globalCN.png', plot=globalCN.plot + labs(title=''), width=16, height=4.6, units='in', dpi=300)
globalCN.plot

## NDBE
nb.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')', sep=''))
globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')
globalBECN.plot = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") 
ggsave(filename='plots/globalBECN.png', plot=globalBECN.plot + labs(title=''), width=16, height=4.6, units='in', dpi=300)
globalBECN.plot

@@

```{r noLGD, echo=F, warning=F, message=F, fig.width=6, fig.height=6, eval=T}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

@@

chh = merge(chh,globalCNdf %>% group_by(chr) %>% dplyr::summarise(minV=min(mean.value), maxV=max(mean.value)),by='chr')
chh = chh %>% dplyr::mutate( ymax = ifelse(coef > 0, maxV, minV) )
chh[c('maxV','minV')] = NULL

chh$gl = factor(ifelse(chh$coef > 0, 'gain', 'loss'))

@@

ggsave(filename='plots/auc/all_roc.png', plot=roc.plot(roc), width=6, height=4, units='in', dpi=300)

@@

ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=2.7, height=3.7, units="in", limitsize=F, dpi = 300, scale=2)

@@

ggsave(filename='plots/pred_calib.png', plot=pcal, width=2.2, height=2.2, units='in', dpi=300, scale=2)

@@

unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
  #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
unadjRR
ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=5.5, height=4.3, units='in', dpi=300)

@@

### Mountain plots using spatial info

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=20}
min.theme = theme(text=element_text(size=12), panel.background=element_blank(), strip.background=element_rect(fill="grey97"), strip.text.y = element_text(size=6),
                   strip.text.x = element_text(size=12), axis.text = element_blank(), axis.ticks = element_blank(),
                   axis.line=element_line(color='black'), panel.grid.major.y =element_line(color='grey95'), panel.grid.major.x = element_blank(),
                   panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines'), panel.spacing.y = unit(0.1, 'lines')   )

mtn.spatial<-function(df, b, limits, title='', pal=NULL) {
  if (is.null(pal))
    pal = RColorBrewer::brewer.pal(length(limits), 'Set2')

  df = merge(df, b[,c('Samplename','Endoscopy.Year','months.before.final','Block','ogj')], by.x='variable',by.y='Samplename')
  df$ogj = factor(df$ogj, ordered = T)
  df$ogj = factor(df$ogj, levels=rev(levels(df$ogj)), ordered = T)
  #pal = pal[which(limits %in% levels(df$nl))]

  altcolor = c(RColorBrewer::brewer.pal(9, 'Greys')[2],RColorBrewer::brewer.pal(9, 'Greys')[4])
  fills = c()
  for (yr in sort(unique(df$Endoscopy.Year))) {
    ac = (which(sort(unique(df$Endoscopy.Year)) == yr) %% 2) + 1
    fills = c(fills, rep(altcolor[ac], length(which(table(subset(df, Endoscopy.Year == yr)$ogj) > 0))))
  }

  blanklabel <- function(s) {
    return('')
  }

  p = ggplot(df, aes(x=chr.length)) + facet_grid(Endoscopy.Year+ogj~chr, scales='free', space='free_x', labeller = labeller(.multi_line = F)) +
      geom_rect(aes(xmin=start, xmax=end, ymin=0, ymax=mean.value, fill=ogj), show.legend=T) + 
      scale_fill_manual(values=pal, limits=limits,  name='OGJ Dist.') +
      labs(title=title, x='') + min.theme

  legend = get.legend(p)
  p = p + theme(legend.position = 'none')

  g = ggplot_gtable(ggplot_build(p))
  strips = grep('strip-r', g$layout$name)

  for (n in 1:length(strips)) {
    i = strips[n]
    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[n]
  }

  #grid.arrange(g)
  return(list('grob'=g, 'legend'=legend))
  #grid::grid.draw(g)  
  #return(g)
}

back$ogj = back$nl
back$ogj[which(back$nl > 0 & back$nl < 3)] = 1 
back$ogj[which(back$nl > 3 & back$nl < 7)] = 2 
back$ogj[which(back$nl >= 7)] = 3

ogjLimits = levels(factor(back$ogj, ordered = T))

df = subset(globalCNdf, Patient == 20 & seg.type != 'arms')
#df = subset(globalCNdf, Patient == 51 & seg.type != 'arms')

# early/late samples
#df = subset(globalCNdf, Patient == pt)
df = merge(globalCNdf, back[,c('Samplename','Endoscopy.Year', 'ogj', 'months.before.final', 'OR','Prediction')], by.x='variable', by.y='Samplename')
df = arrange(df, chr, start, Endoscopy.Year)

## count arms differently than segs??
df2 = df %>% group_by(Status, Patient, Endoscopy.Year) %>% summarise( 
  'max.Pred'=max(Prediction),
  'max.RR'=max(OR),
  'max.Path'=max(Pathology),
  'CNA'=length(which(seg.type != 'arms' & cn %in% c('loss','gain'))),
  'arms'=length(which(seg.type == 'arms' & cn %in% c('loss','gain'))),
  'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
  )
#as.data.frame(df2)
#df2 = subset(df2, max.Path %in% c('BE','ID','LGD'))

df2 = df2 %>% group_by(Status, Patient) %>% mutate(
  'early'= Endoscopy.Year == min(Endoscopy.Year),
  'late' = Endoscopy.Year == max(Endoscopy.Year)
)
df2 = subset(df2, Patient %in% names(which(table(df2$Patient) > 1)))

el = subset(df2, early | late)
el2 = el %>% group_by(Status, Patient) %>% summarise( diff(arms), diff(CNA) )
wtA = round(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value, 3)
if (wtA < 0.0001) wtA = '< 0.0001'
#wilcox.test(subset(el2, Status == 'Progressor')$`diff(CNA)`, subset(el2, Status == 'Non-Progressor')$`diff(CNA)`)

p = ggplot(subset(df2, early|late), aes(Endoscopy.Year, arms, color=Patient)) + facet_grid(~Status) +
  geom_point() +  geom_line(data=subset(df2, early | late)) + theme(legend.position = 'none', text=element_text(size=11)) + 
  plot.theme + labs(x='Endoscopy Year', y='CN arm values', title='Initial vs final endoscopy', subtitle=paste('p', wtA))
p
ggsave(filename='plots/status_hetero.png', plot=p, width=8, height=6, units='in', dpi=300)

wtA = round(wilcox.test(subset(df2, early &  Status == 'Progressor')$arms, subset(df2, early & Status == 'Non-Progressor')$arms)$p.value, 3)
wtCN = round(wilcox.test(subset(df2, early & Status == 'Progressor')$CNA, subset(df2, early & Status == 'Non-Progressor')$CNA)$p.value, 2)

m = melt(subset(df2, early), measure.vars=c('CNA','arms'), id.vars=c('Status'))
p = ggplot(m, aes(Status, value, fill=Status)) + facet_wrap(~variable, scales='free_y') + 
  geom_boxplot(outlier.colour = NA) + geom_jitter(width=0.2) + plot.theme + theme(legend.position = 'none', text=element_text(size=11)) +
  labs(y='', x='', title='CN values, initial endoscopy', subtitle=paste('Arms p=',wtA, '   CNA p=',wtCN, sep=''))
p
ggsave(filename='plots/hetero_initial.png', plot=p, width=6, height=4, units='in', dpi=300)

p2 = mtn.spatial(subset(globalCNdf, Patient == 20 & seg.type != 'arms'), back, ogjLimits,'Patient 20 (P)' )
p6 = mtn.spatial(subset(globalCNdf, Patient == 51 & seg.type != 'arms'), back,ogjLimits,'Patient 51 (NP)' )

# for (i in 1:nrow(sum.patient.data)) {
#   print(i)
#   pt = sum.patient.data[i, 'Patient']
#   status = sum.patient.data[i, 'Status']
# 
#   pms = mtn.spatial(subset(globalCNdf, Patient == pt & seg.type != 'arms'), back, ogjLimits, paste('Patient', pt, '(', status, ')' ))
#   
#   fl = paste('patient_', pt, '.png', sep='')
#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=14, height=6, units='in', dpi=300)
# }

back$ogj = NULL

```{r, echo=F, fig.width=2, fig.height=4}

grid.arrange(p2$legend)

ggsave(filename='plots/spatial/patient_20.png', plot=grid.arrange(p2$grob), width=14, height=6, units='in', dpi=300)
ggsave(filename='plots/spatial/patient_51.png', plot=grid.arrange(p6$grob), width=14, height=6, units='in', dpi=300)

At each quintile are the percentage of progressor samples that were predicted within that range, and the hypergeometric p-value (Fishers) for the significance within that quintile. Then our predictions are at best a coin flip in the middle, but highly significant at the outer edges. This would also push the prediction for "high" risk up to `r min(subset(pred.confidence, labels == 'High' )$r1)` while "low" risk is below `r max(subset(pred.confidence, labels == 'Low' )$r2)`.

`r pander(as.data.frame(pred.confidence))`

@@

ggsave(filename='plots/pred_by_path.png', plot=rp, width=5.8, height=4.5, units='in', dpi=300, scale=1.5)

```{r p53IHC, echo=F, warning=F, message=F, fig.height=8, fig.width=8}

#p53Path$Risk[p53Path$Risk == 'High'] = 'Aberrant'
#p53Path$Risk[p53Path$Risk == 'Low'] = 'Normal'

#p53Path$p53.Status = factor(p53Path$p53.Status, levels=c('Aberrant', 'Normal'), ordered = F)

p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
p53Path = arrange(p53Path, Pathology, p53.Status)
col = brewer.pal(11, "PRGn")[c(10,2)]
rp = ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=102)) +
  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
rp
ggsave(filename='plots/p53_pred.png', plot=rp, width=4.5, height=3.6, units='in', dpi=300, scale=1.5)

@@

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='orangered4', show.legend = F) + 
  geom_errorbar(color='grey21', width=0.2) +
  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
rp
ggsave(filename='plots/pred_endo.png', plot=rp, width=5.6, height=4.4, units='in', dpi=300, scale=1.5)

#endos$months.before.final = endos$months.before.final*-1
#me = endos %>% group_by(Patient) %>% dplyr::summarise('max.endo' = max(i), 'months' = min(months.before.final))
```

`r pander(as.data.frame(pred.confidence))`

@@

genes = cbind(subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer) & Mutation.Types != 'T', select=c('Gene.Symbol','Genome.Location', colnames(loc))), 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))

genes = merge(genes, globalCNdf %>% group_by(chr) %>% dplyr::summarise( value = min(mean.value)+sd(mean.value) ), by='chr')

@@

ggsave(filename='plots/hazards/chr17.png', plot=g17, width=7.5, height=3.7, units='in', dpi=300, scale=1.5)

@@

```{r echo=F, eval=T}
dirs = c('~/Data/Ellie/Analysis/downsampled5G_BEAdjacent',
         '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent_normal/',
          '~/Data/Ellie/Analysis/downsampled5G_NDBE')

@@

predfiles = sapply(dirs, function(x) {
  list.files(x, 'predictions_all_exAHM0320.Rdata', full.names = T)
})

predObj = list()
for (i in 1:length(predfiles)) {
  if (length(predfiles[[i]]) > 0) {
      load(predfiles[[i]], verbose=T)
      dspred$Type = basename(names(predfiles[i]))
      predObj[[i]] = dspred
      rm(dspred)
  } else {
    files = list.files(names(predfiles[i]), pattern='raw', full.names=T)
    sampleNames = sub('_raw.txt', '', basename(files))
    dspred = data.frame(matrix(ncol=2, nrow=length(sampleNames), dimnames=list(sampleNames,c('Prob','RR'))))
    for (n in 1:length(files)) {
      print(n)
      print(sampleNames[n])

@@

      dspred[n,] = c(preds[,1],RR[,1])
      print(dspred[1:n,])
  }

    dspred$Type = basename(names(predfilels[i]))
    predObj[[i]] = dspred

    save(dspred, file=paste(names(predfiles[i]), 'predictions_all_exAHM0320.Rdata', sep='/'))
    rm(dspred)
  }

  dspred = do.call(rbind, predObj)
